Efficient Optimal Approximation of Discrete Random Variables \\ for Estimation of Probabilities of Missing Deadlines

### 2018-11-12 rebuttal.txt

We thank the reviewers for the constructive and deep feedback. See the responses to all specific issues below. We will also fix the small typos identified by the reviewers.

Reviewer #1:
-----------------
Thank you for the corrections. We will incorporate them in the final version.

Regarding "dual(X,m) -> dual(X, e_m)" and "(m', e) -> (m', e_{i,j}) ". You are correct.

Regarding Algorithm 3, to make the algorithm more readable, we will move the update to a separate line and execute it iff $m' \geq m$. This does not change the complexity and only adds candidates to S (does not remove any candidate from it) thus it does not break the correctness of the algorithm.

Reviewer #2:
-----------------
Regarding interest to the broader AI community:  compression of random variables and of other objects used in AI, such as neural networks, is a broad problem with many applications. We focused on a specific problem in this paper in order to provide full details. We plan to continue with other results towards a complete theory and practices in this field.

In page 7, we will change "random variables of 100" to "random variables with $n=100$". Thank you.

We will insert full bibliographic details as proposed.

Reviewer #3:
-----------------
Cohen et al. (2018) is the state of the art to the best of our knowledge. There is also a recent draft on the arxiv that presents an attempt to tackle the optimal approximation problem that we deal with in this paper by extending the algorithm of Cohen et al. (2018). We compared the results of this algorithm with ours and made sure that they produce random variables with equal approximation quality (same Kolmogorov distance from the original random variable). However, the draft on the archive presents a quadratic algorithm while ours complexity is of O(n logn). We also tested that our algorithm runs faster in practice.

Reviewer #4:
-----------------
Indeed, we agree with the reviewer that this work can be seen as a compression algorithm. We believe our work is relevant to a wider audience and we plan to extend it to compression by other metrics and of other objects.

The following are answers to the "Detailed Comments":

- We will add to Section 1 a forward reference to the linear program described in Section 2. Liat, sec 1

- We have some preliminary work on possible extensions of our work to other distances. We believe that we can find optimal approximations relative to these distances in polynomial time, but not in linear time. We didn't add these to the paper because we don't want to publish half-baked results. We plan to add them to a journal version of the paper in the near future.

- We verified that the <= is correct in line 6 of Algorithm 1. The idea is that we extend the interval [b,e] until $c_{e+1}-c_b$ it is larger than $2\varepsilon$ and $c_{e}-c_b$ is smaller or equal to it.

- We agree that the proof of Algorithm 1 can be more detailed. We will elaborate in the final version. Specifically, we will explain that we are not only showing that the solution found is locally optimum, we are showing that it is a *global* optimum. The proof goes by showing that *any* change to the solution will either give a inferior or same approximation distance.

- We will add an example of execution of Algorithm 1 to explain the rationale.

- We will clarify in the proof of Algorithm 2 that the fact that the optimal error is one of the distances between any two steps of the CDFs is because we seek a random variable whose support is included in the original one. Thank you for pointing out this leap in the proof, we believe that the clarifications will improve the final version.

- We will note in the proof of Algorithm 2 the fact that the distance is monotonic with the support size. Thank you for this advice, we believe that it will significantly help future readers.

- Regarding the question "Why do you need k and k' in algorithm 2 and not just l and r": k' is needed for the termination condition. We tried to write the algorithm without m' and k and we felt that it comes too complex to read.

- Regarding the question about the second call to the dichotomy search: If m is not in the set of optimal solutions we may end up with a variable that is not optimal because it can be improved by using m'' such that $m < m'' < m'$. This issue is solved by a direct application of the dual algorithm with the computed $e_k$. We will elaborate.

- We will add some background and explanation on the use of the sorted matrix.

- Regarding 850s that the linear programming solver consumed: In the specific experiment discussed in this context the algorithms (LP and our algorithm) are both implemented in Wolfram Mathematica. Therefore, the run-time comparison is fair. We will report on the number of variables and constraints. Unfortunately, we don't have hand-on experience with cplex or gurobi. Note that the 850s is because this is not a pure continuous LP problem since it involves counting zeroes and optimizing for the maximum norm.

### 2018-11-12 Kolmogorov_Approx.tex
Updates from ShareLaTeX

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required

\usepackage[activate={true,nocompatibility},final,tracking=true,kerning=true,spacing=true,factor=1100,stretch=10,shrink=100]{microtype}

@@

  \pdfinfo{
	/Title (Optimal Approximation of Discrete Random Variables for Estimation of Probabilities of Missing Deadlines
	)
	/Author ()}
\setcounter{secnumdepth}{0}

\begin{document}

\title{Optimal Approximation of Discrete Random Variables \\ for Estimation of Probabilities of Missing Deadlines}
\author{}
\maketitle
\begin{abstract}
We present an efficient algorithm that, given a discrete random variable $X$ and a number $m$, computes a random variable whose support if of size at most $m$ and whose Kolmogorov distance from $X$ is minimal. We present the algorithm, analyse its correctness and computational complexity, and present a detailed empirical evaluation that shows how it performs in practice. The main application that we examine, which is our motivation for this work, is estimation of the probability missing deadlines in series-parallel schedules. Since exact computation of these probabilities is NP-hard, we propose to use the algorithm described in this paper to obtain an approximation.   
\end{abstract}

@@

Various approaches for approximation of probability distributions are studied in the literature~\cite{PS77,AMCR83,vidyasagar2012metric,cohen2015estimating,pavlikov2016cvar,CohenGW18}. 
These approaches vary in the types random variables considered, how they are represented, and in the criteria used for evaluation of the quality of the approximations. In this paper we propose an approach for compressing the probability mass function of a random variable $X$ such that the errors added to queries such as $Pr(X\leq t)$, for  any $t>0$, is minimal. In other words, we minimise the Kolmogorov distance between the approximation and the original random variable. Our motivation for this is estimation of the probability for missing deadlines, as described, e.g., in Cohen et al.~\cite{cohen2015estimating,CohenGW18} and in~\cite{Kashef18}. Specifically, when $X$ represents the probability distribution of the time to complete some complex schedule and we cannot afford to maintain the full table of its probability mass function, we propose an algorithm for producing a smaller table, whose size can be specified, such that probabilities for missing deadlines are preserved as much as possible.

@@

The case study examined in this paper is the problem of task trees with deadlines~\cite{cohen2015estimating,CohenGW18}. A hierarchical plan is a method for representing problems of automated planning in which the dependency among tasks can be given in the form of networks, here we focus on hierarchical plans represented by task trees. The leaves in a task tree are \emph{primitive} actions (or tasks), and the internal nodes are either \emph{sequence} or \emph{parallel} actions. The plans we deal with are of stochastic nature, and the task duration is described as probability distributions in the leaf nodes. We assume that the distributions are independent but {\em not} necessarily identically distributed and that the random variables are discrete and have a finite support.

A sequence node denotes a series of tasks that should be performed consecutively, whereas a parallel node denotes a set of tasks that begin at the same time. A \emph{valid} plan is one that is fulfilled before some given \emph{deadline}, i.e., its \emph{makespan} is less than or equal to the deadline. The objective in this context is to compute the probability that a given plan is valid, or more formally computing $P(X<T)$, where $X$ is a random variable representing the makespan of the plan and $T$ is the deadline. The problem of finding the probability that a task tree satisfies a deadline is known to be NP-hard~\cite{cohen2015estimating}. In fact, even the problem of summing a set of random variables is NP-hard~\cite{mohring2001scheduling}. This is an example of an explicitly given random variable that we need to estimate deadline missing probabilities for.

@@

The size of a random variable is measured here by the size of its support, the set of possible outcomes, $|X|=|X|=|\{x\colon Pr(X=x) \neq 0\}|$. When probability mass functions are maintained as explicit tables, as done in many implementations of statistical software, the size of the support of the variable is proportional to the amount of memory needed to store it and to the complexity of the computations that manipulate it.

Together, the exact notion of optimality of the approximation targeted in this paper is:
\begin{definition}
	A random variable $X'$ is an optimal $m$-approximation of a random variable $X$ if $|X'| \leq m$ and there is no random variable $X''$ such that $|X''| \leq m$ and $d_K(X,X'') < d_K(X,X')$.
\end{definition}

In these terms, the main contribution of the paper is an efficient (polynomial time and memory) algorithm that takes $X$ and $m$ as parameters and constructs an optimal $m$-approximation of $X$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}\label{sec:relwork}
The most relevant work related to this paper is the papers on approximations of random variables in the context of estimating deadlines~\cite{cohen2015estimating,CohenGW18}. In these papers, $X'$ is defined to be a good approximation of $X$ if $F_{X'}(t) > F_{X}(t)$ for any $t$ and $\sup_t F_{X'}(t) - F_{X}(t)$ is small. Note that this measure is not a proper distance measure because it is not symmetric. The motivation given in these papers for using this type of approximation is for cases where overestimation of the probability of missing a deadline is acceptable but underestimation is not. In Section~\ref{sec:exp}, we consider the same case-studies examined by Cohen et al. and show how the algorithm proposed in this paper performs relative to the algorithms proposed there when both over- and under- estimations are allowed. As expected, the Kolmogorov distance between the approximated and the original random variable is considerably smaller when using the algorithm proposed in this paper.

In the technical level, the problem we study in this paper is similar to the problem of approximating a set of 2-D points by a step function. The study of this problem was motivated by query optimisation and histogram constructions in database management systems~\cite{applf12,applf13,applf14,applf17,applf18,Fournier2011} and computational geometry~\cite{diaz2001fitting,fournier2008fitting}. There are, however, two technically significant differences between the problem studied in the context of databases and the problem we analyse in this paper. The first difference is that in the context of approximation of random variables, the step function (which is the cumulative distribution function in our context)  must end with a value one, since we are dealing with random variables which sums to one. The second difference is that the first step is not counted because there is no need to put a value in the support of the approximated random variable to generate this first step. These cannot be addressed by adding a constant (two) to $m$ because the first step is always present and because the requirement to end with the value one, restricts the set of eligible step functions. See Section~\ref{sec:algEfficient} for more details.

Another relevant prior work is the theory of Sparse Approximation (aka Sparse Representation) that deals with sparse solutions for systems of linear equations, as follows. 
Given a matrix $D \in \mathbb{R}^{n \times p}$ and a vector $x \in \mathbb{R}^n$, the most studied sparse representation problem is finding the sparsest possible representation $\alpha \in \mathbb{R}^p$ satisfying $x = D\alpha$:
\[
\min_{\alpha \in \mathbb{R}^p} \|\alpha\|_0 \text{ subject to } x = D\alpha
\]
where $\|\alpha\|_0 = |\{ i \in [p]: \alpha_i \neq 0 \}|$ is the $\ell_0$ pseudo-norm, counting the number of non-zero coordinates of $\alpha$. This problem is known to be NP-hard with a reduction to NP-complete subset selection problems.
In these terms, using also the $\ell_\infty$ norm that represents the maximal coordinate and the $\ell_1$ norm that represents the sum of the coordinates, our problem can be phrased as:
\[
\min_{\alpha \in [0,\infty)^p}\|x - D\alpha\|_{\infty} \text{ subject to }  \|\alpha\|_0 = m \text{ and } \|\alpha\|_1=1
\]
where $D$ is the lower unitriangular matrix, $x$ is related to $X$ such that the $i$th coordinate of $x$ is $F_X(x_i)$ where $\support(X)=\{x_1 < \cdots < x_n\}$ and $\alpha$ is related to $X'$ such that the $i$th coordinate of $\alpha$ is $f_{X'}(x_i)$. The functions $F_X$ and $f_{X'}$ represent, respectively, the cumulative distribution function of $X$ and the mass distribution function of $X'$, i.e.,  the coordinates of $x$ are positive and monotonically increasing and its last coordinate is one. We show that this specific sparse representation problem can be solved in $O(n^2m)$ time and $O(m^2)$ memory.

%The presented work is also related to the research on binning in statistical inference. Consider, for example, the problem of credit scoring~\cite{zeng2017comparison} that deals with separating good applicants from bad applicants where the Kolmogorov–Smirnov statistic KS is a standard measure. The KS comparison is often preceded by a procedure called binning where small values in the the probability mass function are moved to nearby values. There are many methods for binning~\cite{mays2001handbook,refaat2011credit,bolton2010logistic,siddiqi2012credit}.
%In this context, our algorithm can be considered as a binning strategy that provides optimality guarantees with respect to the Kolmogorov distance.

@@

A problem very similar to ours is termed ``order reduction'' by Vidyasagar in~\cite{vidyasagar2012metric}. There, the author defines an information-theoretic based distance between discrete random variables and studies the problem of finding a variable whose support is of size $m$ and its distance from $X$ is as small as possible (where $X$ and $m$ are given). The only difference between this and the problem studied in this paper, is that Vidyasagar examines a different notion of distance. Vidyasagar proves that computing the distance (that he considers) between two probability distributions, and computing the optimal reduced order approximation, are both NP-hard problems, because they can both be reduced to nonstandard bin-packing problems. He then develops efficient greedy approximation algorithms. In contrast, our study shows that there are efficient solutions to these problems when the Kolmogorov distance is considered.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algorithms for optimal approximation}\label{sec:algEfficient}
%In the scope of this section, let $X$ be a random variable ket $m$ be such that  $0<m\leq |X|$. We present algorithms for finding optimal $m$-approximations of $X$.

We begin with presenting an algorithm for solving a problem that is dual to the $m$-approximation problem.  The dual problem is: given a random variable $X$ and $0 \leq \varepsilon \leq 1$, find a new random variable $X'$ such that $d_k(X,X')\leq \varepsilon$ and $|X'|$ is minimal.

We assume that the input to Algorithm~\ref{alg:dual} that solves the dual problem is a representation of the variable $X$ as a sorted CDF, i.e.,  as an array $X_{CDF}=\{(x_i, c_i)\}_{i=1}^n$ such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$. In lines 2 we perform a single pass over $X_{CDF}$ starting from index 1 to find the last index where the cumulative distribution function of $X$ is less than $\varepsilon$, this index is now $f$. In line 3 we start from the last index of $X_{CDF}$ which is $n$, and in decreasing order seek for the last index where the cumulative distribution function of $X$ is less than $1-\varepsilon$, this index is now $l$. The last part of the algorithm is a pass on all elements in $X_{CDF}$ between the indices $f$ and $l$ in order to construct the $m$-approximation random variable of $X$, $X'$. In lines 6-7 we check if the difference between the first element in the current set of indices (which we call $b$) and the last element in the current set of indices (which we call $e+1$) is less than $2\varepsilon$. If yes, we add the pair $\{(x_{b}, (c_b+c_e)/2 - s) \}$ to the set $S$. Otherwise, we add one to the counter of $e$.

\begin{algorithm}
	\DontPrintSemicolon
	%Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$.\;
    $f \gets 0$, $l \gets n$+1\; 
    \lWhile{$c_{f +1} \leq \varepsilon$}{$f \gets f + 1$}

@@

	$S \gets \emptyset$, $s \gets 0$,$b \gets f$, $e \gets f$\;
	\While{$e<l$}{
	    \While{$c_{e+1}-c_{b} \leq 2\varepsilon \wedge e+1 <l$}{$e \gets e+1$}
	    $S \gets S \cup \{(x_{b}, (c_b+c_e)/2 - s) \}$\;
        $s \gets  (c_b+c_e)/2$, $b \gets e+1$\;
	}
	$S \gets S \cup \{(x_{l}, 1 - s) \}$\;
    \Return{A r.v. $X'$ such that $Pr(X'{=}x)=c$ if there is $c$ such that $(x, c) \in S$ and $Pr(X'{=}x)=0$ otherwise.}

@@

\SetKwRepeat{Do}{do}{while}
%which c? why -s? name without any meaning, cant understand

\begin{theorem}\label{the:correctnessDual}
	If  $\{(x_i, c_i)\}_{i=1}^n$ is such that $c_i=Pr(X \leq x_i)$ and $\supp(X)=\{x_1 < \cdots < x_n\}$ then  
    $$dual(\{(x_i, c_i)\}_{i=1}^n,\varepsilon) \in \underset{X'\in \bar{\mathcal{B}}(X,\varepsilon)}{\operatorname{arg\,min}}\, |X'|$$  where $\bar{\mathcal{B}}(X,\varepsilon)=\{X'\colon d_K(X,X')\leq \varepsilon\}$.
\end{theorem}
\begin{proof}
The number of level sets of the CDF of $X'$ is minimal: (1) The first and the last sets are of maximal length; (2) By construction, an extension of any of the other sets to the right will generate a random variable whose Kolmogorov distance from $X$ is bigger than $\varepsilon$; and (3) Extension of a level set to the left may either leave the number of level sets unchanged or combine it with the previous set, which will enlarge the Kolmogorov distance because it is equivalent to extending the previous set to the right.
\end{proof}

\begin{theorem}\label{the:complexityDual}
	$dual(\{(x_i, c_i)\}_{i=1}^n,\varepsilon)$ runs in time $O(n)$, using $O(n)$ memory.
\end{theorem}
\begin{proof}
This algorithm describes a single pass over $\{(x_i, c_i)\}_{i=1}^n$. Line 2 and line 3 are easy to follow, each takes $O(n)$ in the worst case. Lines 5-9 also describe a single pass since the counter $e$ is updated to $e+1$ at most $n$ times. All together we get run-time complexity of $O(n)$. We are constructing the set $S$ which is of size $n$ in the worst case, therefore, memory complexity is also $O(n)$. 
\end{proof}

>>>>

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
\usepackage{float}

\usepackage[activate={true,nocompatibility},final,tracking=true,kerning=true,spacing=true,factor=1100,stretch=10,shrink=40]{microtype}

\makeatletter

% this creates a custom and simpler ruled box style
\newcommand\floatc@simplerule[2]{{\@fs@cfont #1 #2}\par}
\newcommand\fs@simplerule{\def\@fs@cfont{\bfseries}\let\@fs@capt\floatc@simplerule
  \def\@fs@pre{\hrule height.8pt depth0pt \kern4pt}%
  \def\@fs@post{\kern4pt\hrule height.8pt depth0pt \kern4pt \relax}%
  \def\@fs@mid{\kern8pt}%
  \let\@fs@iftopcapt\iftrue}

% this code block defines the new and custom floatbox float environment
\floatstyle{simplerule}
\newfloat{exmpl}{thp}{lob}[section]
\floatname{exmpl}{Example}

@@

  \pdfinfo{
	/Title (Efficient Optimal Approximation of Discrete Random Variables for Estimation of Probabilities of Missing Deadlines
	)
	/Author (Liat Cohen, Gera Weiss)}

\setcounter{secnumdepth}{1}  
\pgfplotsset{compat=1.5}
\begin{document}

\title{Efficient Optimal Approximation of Discrete Random Variables \\ for Estimation of Probabilities of Missing Deadlines}
\author{Liat Cohen \and Gera Weiss\\
    Department of Computer Science\\
		Ben-Gurion University of the Negev, Beer-Sheva 84105, Israel\\
		\{liati,geraw\}@cs.bgu.ac.il\\}
\maketitle
\begin{abstract}
We present an efficient algorithm that, given a discrete random variable $X$ and a number $m$, computes a random variable whose support is of size at most $m$ and whose Kolmogorov distance from $X$ is minimal. We present some variants of the algorithm, analyse their correctness and computational complexity, and present a detailed empirical evaluation that shows how they performs in practice. The main application that we examine, which is our motivation for this work, is estimation of the probability missing deadlines in series-parallel schedules. Since exact computation of these probabilities is NP-hard, we propose to use the algorithms described in this paper to obtain an approximation.   
\end{abstract}

@@

Various approaches for approximation of probability distributions are studied in the literature~\cite{PS77,AMCR83,vidyasagar2012metric,cohen2015estimating,pavlikov2016cvar,CohenGW18}. 
These approaches vary in the types random variables considered, how they are represented, and in the criteria used for evaluation of the quality of the approximations. In this paper we propose an approach for compressing the probability mass function of a random variable $X$ such that the errors added to queries such as $Pr(X\leq t)$, for  any $t>0$, is minimal. In other words, we minimise the Kolmogorov distance between the approximation and the original variable, see alternative definition in Equation~\eqref{eq:linear}.

Our main motivation for this work is estimation of the probability for missing deadlines, as described, e.g., in Cohen et al.~\cite{cohen2015estimating,CohenGW18} and in~\cite{Kashef18}. Specifically, when $X$ represents the probability distribution of the time to complete some complex schedule and we cannot afford to maintain the full table of its probability mass function, we propose an algorithm for producing a smaller table, whose size can be specified, such that probabilities for missing deadlines are preserved as much as possible.

@@

%The case study examined in this paper is the problem of task trees with deadlines~\cite{cohen2015estimating,CohenGW18}. A hierarchical plan is a method for representing problems of automated planning in which the dependency among tasks can be given in the form of networks, here we focus on hierarchical plans represented by task trees. The leaves in a task tree are \emph{primitive} actions (or tasks), and the internal nodes are either \emph{sequence} or \emph{parallel} actions. The plans we deal with are of stochastic nature, and the task duration is described as probability distributions in the leaf nodes. We assume that the distributions are independent but {\em not} necessarily identically distributed and that the random variables are discrete and have a finite support.

%A sequence node denotes a series of tasks that should be performed consecutively, whereas a parallel node denotes a set of tasks that begin at the same time. A \emph{valid} plan is one that is fulfilled before some given \emph{deadline}, i.e., its \emph{makespan} is less than or equal to the deadline. The objective in this context is to compute the probability that a given plan is valid, or more formally computing $P(X<T)$, where $X$ is a random variable representing the makespan of the plan and $T$ is the deadline. The problem of finding the probability that a task tree satisfies a deadline is known to be NP-hard~\cite{cohen2015estimating}. In fact, even the problem of summing a set of random variables is NP-hard~\cite{mohring2001scheduling}. This is an example of an explicitly given random variable that we need to estimate deadline missing probabilities for.

@@

The size of a random variable is measured by the size of its support, the set of possible outcomes, $|X|{=}|\{x\colon Pr(X{=}x) \neq 0\}|$. When probability mass functions are maintained as tables, as done in many implementations of statistical software, the support size is proportional to the memory needed to store the variable and to the complexity of the computations that manipulate it. The exact notion of optimality of the approximation targeted in this paper is:

\begin{definition}
	A random variable $X'$ is an optimal $m$-approximation of a random variable $X$ if $|X'| \leq m$ and there is no random variable $X''$ such that $|X''| \leq m$ and $d_K(X,X'') < d_K(X,X')$.
\end{definition}

In these terms, the main contribution of the paper is an efficient (linear time and constant memory) algorithm that takes $X$ and $m$ as parameters and constructs an optimal $m$-approximation of $X$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}\label{sec:relwork}
The most relevant work related to this paper is the papers on approximations of random variables in the context of estimating deadlines~\cite{cohen2015estimating,CohenGW18}. In these papers, $X'$ is defined to be a good approximation of $X$ if $F_{X'}(t) > F_{X}(t)$ for any $t$ and $\sup_t F_{X'}(t) - F_{X}(t)$ is small. Note that this measure is not a proper distance measure because it is not symmetric. The motivation given in these papers for using this type of approximation is for cases where overestimation of the probability of missing a deadline is acceptable but underestimation is not. We consider in this paper the same case-studies examined by Cohen et al. and show how the algorithm proposed in this paper performs relative to the algorithms proposed there when both over- and under- estimations are allowed. As expected, the Kolmogorov distance between the approximated and the original random variable is considerably smaller when using the algorithm proposed in this paper.

In the technical level, the problem we study in this paper is similar to the problem of approximating a set of 2-D points by a step function. The study of this problem was motivated by query optimisation and histogram constructions in database management systems~\cite{applf12,applf13,applf14,applf17,applf18,Fournier2011} and computational geometry~\cite{diaz2001fitting,fournier2008fitting}. There are, however, two technically significant differences between the problem studied in the context of databases and the problem we analyse in this paper. The first difference is that in the context of approximation of random variables, the step function (which is the cumulative distribution function in our context)  must end with a value one, since we are dealing with random variables which sums to one. The second difference is that the first step is not counted because there is no need to put a value in the support of the approximated random variable to generate this first step. These cannot be addressed by adding a constant (two) to $m$ because the first step is always present and because the requirement to end with the value one, restricts the set of eligible step functions.

Another relevant prior work is the theory of Sparse Approximation (aka Sparse Representation) that deals with sparse solutions for systems of linear equations, as follows. 
Given a matrix $D \in \mathbb{R}^{n \times p}$ and a vector $x \in \mathbb{R}^n$, the most studied sparse representation problem is finding 
%the sparsest possible representation $\alpha \in \mathbb{R}^p$ satisfying $x = D\alpha$:
$$
\min_{\alpha \in \mathbb{R}^p} \|\alpha\|_0 \text{ subject to } x = D\alpha
$$
where $\|\alpha\|_0 = |\{ i \in [p]: \alpha_i \neq 0 \}|$ is the $\ell_0$ pseudo-norm, counting the number of non-zero coordinates of $\alpha$. This problem is known to be NP-hard with a reduction to NP-complete subset selection problems.
In these terms, using also the $\ell_\infty$ norm that represents the maximal coordinate and the $\ell_1$ norm that represents the sum of the coordinates, our problem can be phrased as:
\begin{equation}\label{eq:linear}
\min_{\alpha \in [0,\infty)^p}\|x - D\alpha\|_{\infty} \text{ subject to }  \|\alpha\|_0 = m \text{ and } \|\alpha\|_1=1
\end{equation}
where $D$ is the lower unitriangular matrix, $x$ is related to $X$ such that the $i$th coordinate of $x$ is $F_X(x_i)$ where $\support(X)=\{x_1 < \cdots < x_n\}$ and $\alpha$ is related to $X'$ such that the $i$th coordinate of $\alpha$ is $f_{X'}(x_i)$. The functions $F_X$ and $f_{X'}$ represent, respectively, the cumulative distribution function of $X$ and the mass distribution function of $X'$, i.e.,  the coordinates of $x$ are positive and monotonically increasing and its last coordinate is one. %We show that this specific sparse representation problem can be solved in $O(n^2m)$ time and $O(m^2)$ memory.

The presented work is also related to the research on binning in statistical inference. Consider, for example, the problem of credit scoring~\cite{zeng2017comparison} that deals with separating good applicants from bad applicants where the Kolmogorov–Smirnov statistic KS is a standard measure. The KS comparison is often preceded by a procedure called binning where small values in the probability mass function are moved to nearby values. There are many methods for binning~\cite{mays2001handbook,refaat2011credit,bolton2010logistic,siddiqi2012credit}.
In this context, our algorithm can be considered as a binning strategy that provides optimality guarantees with respect to the Kolmogorov distance.

@@

A problem very similar to ours is termed ``order reduction'' by Vidyasagar in~\cite{vidyasagar2012metric}. There, the author defines an information-theoretic based distance between discrete random variables and studies the problem of finding a variable whose support is of size $m$ and its distance from $X$ is as small as possible (where $X$ and $m$ are given). The main difference between this and the problem studied in this paper, is that Vidyasagar examines a different notion of distance. Vidyasagar proves that computing the distance (that he considers) between two probability distributions, and computing the optimal reduced order approximation, are both NP-hard problems, because they can both be reduced to nonstandard bin-packing problems. He then develops efficient greedy approximation algorithms. In contrast, our study shows that there are efficient solutions to these problems when the Kolmogorov distance is considered.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algorithms for optimal approximation}\label{sec:algEfficient}
We begin with presenting an algorithm for solving a problem that is dual to the $m$-approximation problem: given a random variable $X$ and $0 \leq \varepsilon \leq 1$, find a new random variable $X'$ such that $d_k(X,X')\leq \varepsilon$ and $|X'|$ is minimal.

We assume that the input to Algorithm~\ref{alg:dual} that solves the dual problem is a representation of the variable $X$ as a sorted CDF, i.e.,  as an array $X_{CDF}=\{(x_i, c_i)\}_{i=1}^n$ such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$. In line~2 we perform a single pass over $X_{CDF}$ starting from the first index to find the last index where the cumulative distribution function of $X$ is less than $\varepsilon$, this index is stored in the variable $f$. In line 3 we start from the last index of $X_{CDF}$ which is $n$, and seek in decreasing order for the last index where the cumulative distribution function of $X$ is less than $1-\varepsilon$. This index is stored at the variable $l$. The last part of the algorithm is a pass on all elements of $X_{CDF}$ between the indices $f$ and $l$ in order to construct the random variable $X'$. In lines 6-7 we check if the difference between the first element in the current set of indices (which we call $b$) and the last element in the current set of indices (which we call $e+1$) is less than $2\varepsilon$. If yes, we add the pair $(x_{b}, (c_b+c_e)/2 - s)$ to the set $S$. Otherwise, we add one to $e$. See running Example~\ref{exmpl:dual}.

\begin{algorithm}
	\DontPrintSemicolon
	%Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$.\;
    $f \gets 1$, $l \gets n$+1\; 
    \lWhile{$c_{f} \leq \varepsilon$}{$f \gets f + 1$}

@@

	$S \gets \emptyset$, $s \gets 0$,$b \gets f$, $e \gets f$\;
	\While{$e<l$}{
	    \While{$c_{e+1}-c_{b} \leq 2\varepsilon \wedge e <l$}{$e \gets e+1$}
	    $S \gets S \cup \{(x_{b}, (c_b+c_e)/2 - s) \}$\;
        $s \gets  (c_b+c_e)/2$, $b \gets e\gets e+1$\;
	}
	$S \gets S \cup \{(x_{l}, 1 - s) \}$\;
    \Return{A r.v. $X'$ such that $Pr(X'{=}x)=c$ if there is $c$ such that $(x, c) \in S$ and $Pr(X'{=}x)=0$ otherwise.}

@@

\SetKwRepeat{Do}{do}{while}

\begin{example}
\label{exmpl:dual}
When $dual$ is invoked with the parameters  $X{=}\{(1, 0.3), (2, 0.7), (3, 0.9), (4,1)\}$ and $\varepsilon{=}0.1$: Lines 2 and 3 set $f=1$ and $l=4$. After an iteration of the main loop (line 5), $S = \{(1,0.3)\}$, $s=0.3$, and $b=e=2$. After a second iteration, $S = \{(1,0.3), (2, 0.5)\}$, $s=0.8$, and $b=e=4$. At the end, $S = \{(1,0.3), (2, 0.5), (4, 0.2)\}$.
\end{example}

\begin{proposition}\label{the:correctnessDual}
	If  $\{(x_i, c_i)\}_{i=1}^n$ is such that $c_i=Pr(X \leq x_i)$ and $\supp(X)=\{x_1 < \cdots < x_n\}$ then  
    $$dual(\{(x_i, c_i)\}_{i=1}^n,\varepsilon) \in \underset{X'\in \bar{\mathcal{B}}(X,\varepsilon)}{\operatorname{arg\,min}}\, |X'|$$  where $\bar{\mathcal{B}}(X,\varepsilon)=\{X'\colon d_K(X,X')\leq \varepsilon\}$.
\end{proposition}
\begin{proof}
The number of level sets of the CDF of $X'$ is minimal: (1) The first and the last sets are of maximal length; (2) By construction, an extension of any of the other sets to the right will generate a random variable whose Kolmogorov distance from $X$ is bigger than $\varepsilon$; and (3) Extension of a level set to the left may either leave the number of level sets unchanged or combine it with the previous set, which will enlarge the Kolmogorov distance because it is equivalent to extending the previous set to the right. Thus, there is no random variable whose Kolmogorov distance from $X$ is smaller or equal to $\varepsilon$ and its support is smaller than $|X'|$.
\end{proof}

\begin{proposition}\label{the:complexityDual}
	$dual(\{(x_i, c_i)\}_{i=1}^n,\varepsilon)$ runs in time $O(n)$, using $O(n)$ memory.
\end{proposition}
\begin{proof}
This algorithm describes a single pass over $\{(x_i, c_i)\}_{i=1}^n$. Lines 2 and 3 are easy to follow, each takes $O(n)$ in the worst case. Lines 5-9 also describe a single pass since the counter $e$ is updated to $e+1$ at most $n$ times. All together we get run-time complexity of $O(n)$. We are constructing the set $S$ which is of size $n$ in the worst case, therefore, memory complexity is $O(n)$. 
\end{proof}

The first solution to the $m$-approximation problem we present is the $binsApprox(X,m)$ algorithm which is based on~\cite{diaz2001fitting}. There are couple of significant changes between our $binsApprox(X,m)$ algorithm and the algorithm suggested by D{\'i}az-B{\'a}nez and Mesa~\shortcite{diaz2001fitting} addressing the differences presented in the Related work section. The algorithm $binsApprox(X,m)$ gets as input a random variable $X$ and some number $m$. Again, we do not mind the original representation of $X$ since we can transform it to a sorted CDF representation in $O(nlog(n))$ run-time as in line 1 which we call $X_{CDF}$. In lines 3-8 we compute all possible errors, in other words, all possible $d_{K}(X,X')$ such that $X'$ is an approximation of $X$ and $\support(X') \subseteq \support(X)$. The error is just the difference between the CDF values of every two elements in $X_{CDF}$. After computing the set $E$, we sort it (line 9) in order to perform a binary search. In lines 10-19 we perform a binary search over the set $E$, in every step of the binary search we run the $X' = dual(X,\varepsilon)$ algorithm. If the size of $|X'|>m$ then we know that the error $\varepsilon$ is too small and need to extend the search in the right side of $E$, if the size of $|X'|<m$ then we know that the error $\varepsilon$ is too big and need to extend the search in the left side of $E$, otherwise we found the correct $m$. 
We run the search twice to handle the extreme case where $m \notin \{|dual(X,\varepsilon)| \colon \varepsilon\in E \}$. In this case we may end up with a variable that is not optimal because it can be improved by using $m''$ such that $m < m'' < m'$. We find the optimal $m''$  by the second round ($i=2$) that run after line 18 that sets $m$ to be the $m'$ found in the first round.
